Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-16 nips_2018.tex
example task program and DSL at beginning

@@

%% Automatically inducing programs from examples is a long-standing goal
%% of artificial intelligence.  Recent work in program induction show
%% that deep networks can be very successful at learning to write certain
%% kinds of programs (e.g., RobustFill:~\cite{devlin2017robustfill} and
%% DeepCoder:~\cite{balog2016deepcoder}), as can symbolic search
%% techniques (e.g., Metagol:~\cite{muggleton2015meta},
%% FlashFill:~\cite{gulwani2011automating}, and Sketch~\cite{solar2008program}).
%% However, the success of these approaches -- both symbolic and neural --
%% relies crucially upon a hand-engineered \emph{Domain Specific Language} (DSL).
%% The DSL is an inventory of restricted programming primitives,
%% and imparts domain specific knowledge about the space of programs.
%% To what extent can we dispense with hand-engineered DSLs?

%% This work proposes \emph{learning} the DSL.
%% We consider the setting where we have a collection of related programming tasks,
%% each specified by a set of input/output examples.
%% We do not assume that the tasks are annotated with
%% ground-truth programs.
%% Our algorithm, called \systemEnding,
%% does three things:
%% it finds programs that solve each of the tasks;
%% it creates its own DSL by discovering and reusing
%% domain-specific subroutines;
%% and it learns a neural network that
%% efficiently searches for programs in the DSL.
%% The purpose of the DSL
%% is to expose the commonalities across the programs that solve tasks within a domain,
%% letting the agent solve other related program learning problems.

%% starting with a Lisp-like representation.
%% We introduce an algorithm, called the \system,
%% which creates its own DSL by discovering and then reusing
%% new useful programming idioms and subroutines.

%% Imagine an agent faced with a new domain of unfamiliar tasks: think
%% playing new kinds of videogames, authoring Python code, or navigating
%% through mazes.  Our agent has at its disposal a basis of primitive
%% actions it can compose to propose solutions to these tasks, but it is
%% no idea what kinds of primitives are appropriate for which tasks
%% nor does it know the higher-level vocabulary in which solutions are
%% best expressed.  How can our agent learn to solve problems in this new
%% domain?

%% The AI and machine learning literature contains two broad takes on this problem.
%% The first take is that the agent should come up with a better representation of the space of solutions,
%% for example, by inventing new primitive actions: see \emph{options} in reinforcement learning~\cite{stolle2002learning}, the EC algorithm in program synthesis~\cite{Dechter:2013:BLV:2540128.2540316}, or predicate invention in inductive logic programming~\cite{muggleton2015meta}.
%% The second take is that the agent should learn a discriminative model mapping problems to a distribution over solutions: for example, policy gradient methods in reinforcement learning or neural models of program synthesis~\cite{devlin2017robustfill,balog2016deepcoder}.
%% Our contribution is a general algorithm for fusing these two takes on the problem:
%% we propose jointly inducing a representation language, called a \emph{Domain Specific Language} (DSL),
%% alongside a bottom-up discriminative model that regresses from problems to solutions.
%% We evaluate our algorithm on four domains:
%% building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and functions on lists.
%% %Because we want a domain-general
%% %approach,
%% We model solutions to tasks as
%% programs,
%% which frames our problem as one of program induction.

%% \begin{figure}
%%   \centering

%%   \begin{tikzpicture}

%%   \node[latent] at (0.5,3) (d){$\mathcal{D}$};
%%   \node[latent] at (1.5,3) (t){$\theta$};
%%   \node[latent] at (1,1.75) (z){$p$};
%% %  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
%%   \node[obs] at (1,0.5) (x) {$x$};
%%   \edge {z}{x};
%%   \edge {d,t}{z};
%%   \plate {}{(z)(x)}{$x\in X$};
%%   \node[align = center] at ([yshift = -1.3cm]x.south) {(a) \\generative model};

%%   \begin{scope}[shift = {(0.7,0)}]  
%%   \node[obs] at (3.5,3) (dx){$\mathcal{D}$};
%%   \node[latent] at (3.5,1.9) (zp){$p$};
%%   \node[latent] at (5,1.9) (tx){$\theta^{(x)}$};
%%   \node[obs] at (3.5,0.5) (xp) {$x$};
%%   \plate {}{(zp)(xp)(tx)}{$x\in X$};
%%   \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.south) {(b) \\recognition model};
%%   \draw [->,red] (xp.east) to[out = 0,in = -90] node(nn){} (tx.south);
%%   \draw [->,red] (tx.west) -- (zp.east);
%%   \draw [->,red] (dx.south) -- (zp.north);

%%   \node at (nn) {
%%     \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.15,shift={+(-1,2.5)}}]
%%       \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
%%       \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
%%       \node[rectangle] at (1,1) {};
%%       \foreach \name / \y in {1,...,4}
%%           \node[neuron] (I-\name) at (0,-\y) {};
%%       \foreach \name / \y in {1,...,3}
%%           \node[neuron] (H-\name) at (1,-\y-0.5) {};
%%       \foreach \name / \y in {1,...,4}
%%           \node[neuron] (O-\name) at (2,-\y) {};
%%       \foreach \source in {1,...,4}
%%           \foreach \dest in {1,...,3}
%%               \draw [-latex] (I-\source) -- (H-\dest);
%%       \foreach \source in {1,...,3}
%%           \foreach \dest in {1,...,4}
%%               \draw [-latex] (H-\source) -- (O-\dest);
%%     \end{tikzpicture}
%%   };
%%   \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };

%%   \end{scope}

%%   \end{tikzpicture}
%%   \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives according to probabilities $\theta$. From each program we observe a task $x$ (program inputs/outputs). (b)  Neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to the parameters of a distribution over programs,
%%     $\theta^{(x)}$. Black/red arrows correspond to the generative/recognition models.}\label{graphicalModel}
%% \end{figure}

%% To perform inference, we take inspiration from the Helmholtz machine
%% and the Wake/Sleep algorithm~\cite{hinton1995wake}, which alternates
%% between ``waking'', which draws samples from a recognition model (our
%% $q$); and ``sleeping'', which trains a recognition model on samples
%% from the generative model (our $\mathcal{D},\theta$).  For us, a
%% \textbf{Wake} cycle means using the recognition model to search for
%% programs that solve the tasks.  A sleep cycle means updating these
%% models so that the DSL and recognition model more closely match the
%% distribution of tasks.  An important difference from the Helmholtz
%% machine is that our setup has two distinct sleep phases: a phase where
%% we update the generative model $(\mathcal{D},\theta)$, which we call
%% \textbf{Sleep-G}, and a second phase where we update the recognition
%% model $q$, called \textbf{Sleep-R}.

%% We take the intuitive strategy of exploiting the
%% conditional independence of the programs conditioned on the DSL, and
%% so alternate between estimating the DSL ($\mathcal{D},\theta$) and
%% estimating a posterior distribution over $p$ by searching for
%% programs.  In practice searching for
%% programs is difficult, so we use the recognition model to
%% learn how to search efficiently.

@@

To make this concrete we need to define what programs actually are and
what form $\probability[p |\mathcal{D},\theta]$ takes.
We represent programs as $\lambda$-calculus expressions.
$\lambda$-calculus is a formalism for expressing functional programs
that closely resembles the Lisp programming language.
$\lambda$-calculus includes variables, function application, and the ability to create new functions.
Throughout this paper we will write $\lambda$-calculus expressions in Lisp syntax.
Our programs are all strongly typed.
We use the Hindley-Milner polymorphic typing system~\cite{pierce} which is
used in functional programming languages like OCaml.
%% Type variables are always written using lowercase Greek letters
%% and we write $\alpha\to \beta$ to mean a function that takes an input of type $\alpha$
%% and returns something of type $\beta$.
%% We use the notation $p:\tau$ to mean that the $\lambda$-calculus expression $p$
%% has the type $\tau$.
%% For example, to describe the type of the identity function
%% we would say \code{(lambda (x) x)}$:\alpha\to \alpha$.
%% We say a type $\alpha$ \emph{unifies} with $\tau$ if every expression
%% $p:\alpha$ also satisfies $p:\tau$. Furthermore, the act of \emph{unifying}
%% a type $\alpha$ with $\tau$ is to introduce constraints on the type
%% variables of $\alpha$ to ensure that $\alpha$ unifies with $\tau$.
%% See Supplement for more detail on program representation.
We now define DSLs:

>>>>

\begin{tabular}{r|c|}\cmidrule{2-2}
  Task&
  \begin{tabular}{rcl}
  Nancy FreeHafer&$\longrightarrow$& Dr. Nancy\\
 Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
 Pushmeet Kohli&$\longrightarrow$&???
  \end{tabular}\\\cmidrule{2-2}
  Program&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}\\\cmidrule{2-2}
  DSL&\begin{tabular}{l}
    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y) (cons x y)))}\\
    \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a)}\\
    \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
    \end{tabular}\\\cmidrule{2-2}
  \end{tabular}

@@

@@

To make this concrete we need to define what programs actually are and
what form $\probability[p |\mathcal{D},\theta]$ takes.
We represent programs as $\lambda$-calculus expressions.
$\lambda$-calculus is a formalism for expressing functional programs
that closely resembles the Lisp programming language.
$\lambda$-calculus includes variables, function application, and the ability to create new functions.
Throughout this paper we will write $\lambda$-calculus expressions in Lisp syntax.
Our programs are all strongly typed.
We use the Hindley-Milner polymorphic typing system~\cite{pierce} which is
used in functional programming languages like OCaml and Haskell.
%% Type variables are always written using lowercase Greek letters
%% and we write $\alpha\to \beta$ to mean a function that takes an input of type $\alpha$
%% and returns something of type $\beta$.
%% We use the notation $p:\tau$ to mean that the $\lambda$-calculus expression $p$
%% has the type $\tau$.
%% For example, to describe the type of the identity function
%% we would say \code{(lambda (x) x)}$:\alpha\to \alpha$.
%% We say a type $\alpha$ \emph{unifies} with $\tau$ if every expression
%% $p:\alpha$ also satisfies $p:\tau$. Furthermore, the act of \emph{unifying}
%% a type $\alpha$ with $\tau$ is to introduce constraints on the type
%% variables of $\alpha$ to ensure that $\alpha$ unifies with $\tau$.
%% See Supplement for more detail on program representation.
We now define DSLs:
